---
title: 'Why Trumps Inaugaration Speech Makes Him Unique'
author: "Peter Li (pwl2107)"
date: "9/17/2017"
output: html_document
---
## Overview
In this report, we will explore the inaugaration speechs to understand how Trump is similar and dissimilar to his predecessors and what drives it. 

```{r setup, include=FALSE}
packages.used=c("rvest", "tibble", 
                "sentimentr", "gplots", "dplyr",
                "tm", "syuzhet", "factoextra", 
                "beeswarm", "scales", "RColorBrewer",
                "RANN", "tm", "topicmodels","readtext","tidytext","wordcloud",
                "qdap")

# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}

# load packages
library("qdap")
library("wordcloud")
library("rvest")
library("tibble")
library("sentimentr")
library("gplots")
library("dplyr")
library("tm")
library("syuzhet")
library("factoextra")
library("beeswarm")
library("scales")
library("RColorBrewer")
library("RANN")
library("tm")
library("topicmodels")
library("readtext")
library("tidytext")

source("../lib/speechFuncs.R")
```

## Data Processing
### Read in Speeches

```{r reed data}
folder.path="../data/InauguralSpeeches/"
speeches=list.files(path = folder.path, pattern = "*.txt")
prex.out=substr(speeches, 6, nchar(speeches)-4)

ff.all<-Corpus(DirSource(folder.path))
```

### Basic Text Processing
```{r}
ff.all<-tm_map(ff.all, stripWhitespace)
ff.all<-tm_map(ff.all, content_transformer(tolower))
ff.all<-tm_map(ff.all, removeWords, stopwords("english"))
ff.all<-tm_map(ff.all, removeWords, character(0))
ff.all<-tm_map(ff.all, removePunctuation)

tdm.all<-TermDocumentMatrix(ff.all)

tdm.tidy=tidy(tdm.all)

tdm.overall=summarise(group_by(tdm.tidy, term), sum(count))

tdm.trump=tdm.tidy %>% filter(document == 'inaugDonaldJTrump-1.txt')
tdm.trump=summarise(group_by(tdm.trump, term), sum(count))
```

Just to get a qualitative comparison of Trump's Inaugaration vs. Other Inaugaration, we compare the word clouds. Here is the overall:
```{r}
wordcloud(tdm.overall$term, tdm.overall$`sum(count)`,
          scale=c(5,0.5),
          max.words=100,
          min.freq=1,
          random.order=FALSE,
          rot.per=0.3,
          use.r.layout=T,
          random.color=FALSE,
          colors=brewer.pal(9,"Blues"))
```

And here is Trump:
```{r}
wordcloud(tdm.trump$term, tdm.trump$`sum(count)`,
          scale=c(5,0.5),
          max.words=100,
          min.freq=1,
          random.order=FALSE,
          rot.per=0.3,
          use.r.layout=T,
          random.color=FALSE,
          colors=brewer.pal(9,"Blues"))
```

Because of the fact that overall contains many more documents than just Trump, there is more diversity in the overall word cloud. However, Trump word cloud focuses more heavily on the world Will and not much else, suggesting that his speech was far less focused/diverse than others. There is also less focus on things like justice, freedom, unity, etc. in Trump compared to overall.

### Sentence Text Processing
```{r}
speech.list = readtext("../data/InauguralSpeeches/*")
colnames(speech.list) = c('doc_id','fulltext')

sentence.list=data.frame()
for(i in 1:nrow(speech.list)){
  sentences=sent_detect(speech.list$fulltext[i],
                        endmarks = c("?", ".", "!", "|",";"))
  if(length(sentences)>0){
    emotions=get_nrc_sentiment(sentences)
    word.count=word_count(sentences)
    # colnames(emotions)=paste0("emo.", colnames(emotions))
    # in case the word counts are zeros?
    emotions=diag(1/(word.count+0.01))%*%as.matrix(emotions)
    sentence.list=rbind(sentence.list, 
                        cbind(speech.list[i,-ncol(speech.list)],
                              sentences=as.character(sentences), 
                              word.count,
                              emotions,
                              sent.id=1:length(sentences)
                              )
    )
  }
}

sentence.list=
  sentence.list%>%
  filter(!is.na(word.count)) 

```

## Data Analysis
To analyze the data, we will look at both sentence length and sentiment analysis to compare the speeches.

### Sentence Length
```{r}
sentence.list$word.count = as.numeric(as.character(sentence.list$word.count))
sentence.list.sel=sentence.list 
sentence.list.sel$V1 = substr(sentence.list.sel$V1,6,nchar(as.character(sentence.list.sel$V1))-6)
sentence.list.sel$File=factor(sentence.list.sel$V1)

sentence.list.sel$FileOrdered=reorder(sentence.list.sel$File, 
                                  sentence.list.sel$word.count, 
                                  mean, 
                                  order=T)
par(mar=c(4, 11, 2, 2))

beeswarm(word.count~FileOrdered, 
         data=sentence.list.sel,
         horizontal = TRUE,
         pch=16, col=alpha(brewer.pal(9, "Set1"), 0.6), 
         cex=0.55, cex.axis=0.8, cex.lab=0.8,
         spacing=5/nlevels(sentence.list.sel$FileOrdered),
         las=2, ylab="", xlab="Number of words in a sentence.",
         main="Inaugural Speeches")

sentence.list%>%
  filter(V1=="inaugDonaldJTrump-1.txt",
         word.count<=3) %>%
  select(sentences)
```

And so, we see that compared to the other presidents, Donald Trump's speech tended to have shorter sentences, implying a lower level of sophistication. And upon looking at his short sentences, he repeats the same few short sentences.

### Sentiment Analysis
To conduct sentiment analysis, we will look at sentence length variation throughout the speech and the emotions conveyed in the speech. And for simplicity, we will be visually comparing the last four presidents first term speech (and using the functions from the text mining tutorial).

#### Sentence Length Variation
```{r}
sentence.list$File = substr(sentence.list$V1,6,nchar(as.character(sentence.list$V1))-6)
sentence.list$type = 'inaug'
sentence.list$Term = substr(sentence.list$V1,nchar(as.character(sentence.list$V1))-4,nchar(as.character(sentence.list$V1))-4)

sentence.list$anger = as.numeric(as.character(sentence.list$anger))
sentence.list$anticipation = as.numeric(as.character(sentence.list$anticipation))
sentence.list$disgust = as.numeric(as.character(sentence.list$disgust))
sentence.list$fear = as.numeric(as.character(sentence.list$fear))
sentence.list$joy = as.numeric(as.character(sentence.list$joy))
sentence.list$sadness = as.numeric(as.character(sentence.list$sadness))
sentence.list$surprise = as.numeric(as.character(sentence.list$surprise))
sentence.list$trust = as.numeric(as.character(sentence.list$trust))
sentence.list$negative = as.numeric(as.character(sentence.list$negative))
sentence.list$positive = as.numeric(as.character(sentence.list$positive))

par(mfrow=c(4,1), mar=c(1,0,2,0), bty="n", font.main=1)

f.plotsent.len(In.list=sentence.list, InFile="DonaldJTrump", 
               InType="inaug", InTerm=1, President="Donald Trump")

f.plotsent.len(In.list=sentence.list, InFile="BarackObama", 
               InType="inaug", InTerm=1, President="Barack Obama")

f.plotsent.len(In.list=sentence.list, InFile="GeorgeWBush", 
               InType="inaug", InTerm=1, President="George W. Bush")

f.plotsent.len(In.list=sentence.list, InFile="GeorgeBush", 
               InType="inaug", InTerm=1, President="George Bush")
```

So in terms of sentence length variation, we see the same general trends/volatility.

#### Emotion Analysis
##### Highly Emotional Examples
In terms of highly emotionally sentences, here are some samples from the most recent four presidents first speech:
```{r}
print("Donald Trump")
speech.df=tbl_df(sentence.list)%>%
  filter(File=="DonaldJTrump", type=="inaug", Term==1, word.count>=5)%>%
  select(sentences, anger:trust)
speech.df=as.data.frame(speech.df)
as.character(speech.df$sentences[apply(speech.df[,-1], 2, which.max)])

print("Barack Obama")
speech.df=tbl_df(sentence.list)%>%
  filter(File=="BarackObama", type=="inaug", word.count>=4)%>%
  select(sentences, anger:trust)
speech.df=as.data.frame(speech.df)
as.character(speech.df$sentences[apply(speech.df[,-1], 2, which.max)])

print("George W Bush")
speech.df=tbl_df(sentence.list)%>%
  filter(File=="GeorgeWBush", type=="inaug", Term==1, word.count>=5)%>%
  select(sentences, anger:trust)
speech.df=as.data.frame(speech.df)
as.character(speech.df$sentences[apply(speech.df[,-1], 2, which.max)])

print("George Bush")
speech.df=tbl_df(sentence.list)%>%
  filter(File=="GeorgeBush", type=="inaug", Term==1, word.count>=4)%>%
  select(sentences, anger:trust)
speech.df=as.data.frame(speech.df)
as.character(speech.df$sentences[apply(speech.df[,-1], 2, which.max)])

```
And so, we see that in Trump speech, the main drivers are winning and not having fear, while other presidents focus more on the people (and their well-being), unity, and god.

##### Emotion Distribution
```{r}
emo.means=colMeans(select(sentence.list, anger:trust)>0.01)
barplot(emo.means[order(emo.means)], las=2, col='grey', horiz=T, main="Inaugural Speeches")

tmp = colMeans(select(sentence.list %>% filter(File == 'DonaldJTrump'), anger:trust)>0.01)
tmp[order(tmp)]
```
And so, we see that trumps distribution of emotions tends to follow the same pattern as other presidents (chart is all presidents and the numbers printed after are for Trump)

##### Clustering based on Emotions
```{r}
presid.summary=tbl_df(sentence.list)%>%
  filter(type=="inaug")%>%
  group_by(File)%>%
  summarise(
    anger=mean(anger),
    anticipation=mean(anticipation),
    disgust=mean(disgust),
    fear=mean(fear),
    joy=mean(joy),
    sadness=mean(sadness),
    surprise=mean(surprise),
    trust=mean(trust)
  )

presid.summary=as.data.frame(presid.summary)
rownames(presid.summary)=as.character((presid.summary[,1]))
km.res=kmeans(presid.summary[,-1], iter.max=200,
              5)
fviz_cluster(km.res, 
             stand=F, repel= TRUE,
             data = presid.summary[,-1], xlab="", xaxt="n",
             show.clust.cent=FALSE)

```

And based on the emotion clustering, Trump is grouped in the smallest cluster -- suggesting that his speech is more unique as it is not similar to that many other presidents.

## Conclusion
Overall, Donald Trump's speech is similar to other speechs in terms of structure (sentence length variation and emotion distribution). However, Trump's speech differs greatly in terms of content and sophistication. Trump tends to have shorter sentences, implying lower sophistication. And his content seems very superficial (word cloud shows very little variation in terms of content) and distinct (clustered into group with fewest other presidents).